Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

### 2022-02-08 referee-response.tex

  \textcolor{blue}{
  We now denote $d = p + q$ in Definition 2 and define $\mathcal{X}$ as a subset of $\mathbb{R}^{d}$ such
  that for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
  $x^{\top} I_{p,q} y \in [0,1]$. One simple way to construct such a set
  $\mathcal{X}$ is to first find a collection of $K$ vectors
  $\mathcal{S} = \{\nu_1, \nu_2, \dots, \nu_K\} \subset \mathbb{R}^{d}$
  where $\nu_i^{\top} I_{p,q} \nu_j \in [0,1]$ for all $i,j$ and then define $\mathcal{X}$
  as 
  $$\mathcal{X} = \mathrm{conv}(\mathcal{S}) = \{x = \sum_{i} \lambda_i \nu_i \colon \lambda_i \geq 0 \,\, \text{for all $i$}, \sum_{i} \lambda_i =
  1\}.$$

@@

\item
  p6. Remark 2. What are \(\hat{Z}\), \(V\), and \(D\)?\\
  \textcolor{blue}{
  We had rewrote Remark~2 to more clearly described the matrices
  $\hat{Z}$, $V$ and $D$. To avoid confusion, we now use $\hat{D}$ to
  denote the diagonal matrix whose diagonal entries are the $d = p + q$
  largest eigenvalues (in modulus) of $A$ and we use $\hat{V}$ to denote
  the $n \times d$ matrix whose columns are the corresponding
  eigenvectors. We then use $\hat{Z} = \hat{V} |\hat{D}|^{1/2}$ (where
  the $|\cdot|$ operation is applied elementwise) to denote the
  adjacency spectral embedding (ASE) of $A$ into $\mathbb{R}^{d}$; that is
  to say, the $i$th row of $\hat{Z}$ represents an estimate of the
  latent position $X_i$ (up to some non-identifiability transformation
  $Q \in \mathbb{O}(p,q)$). We also reorganized the paper slightly so
  that the definition and remarks about the non-identifiability of the
  latent positions $X$ in a GRDPG are presented before discussing ASE. }

@@

\item
  p7 line 40-46. I don't understand the interest of this paragraph.\\
  \textcolor{blue}{
  The purpose of the additional material in the proof of Theorem 1 is to
  first present the proof in the special case when $K = 2$ so as to
  build intuition for the general case of $K \geq 2$. 
  More specificially for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
  where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of
  order 2, and the latent vectors lie in the union of two $2$-dimensional
  orthogonal subspaces, i.e., the rows of $X$ consist of
  vectors that lie in $\mathcal{S}_1 \cup \mathcal{S}_2$ where
  $\mathcal{S}_1$ and $\mathcal{S}_2$ are both $2$-dimensional
  subspaces of $\mathbb{R}^{4}$ with $\mathcal{S}_1 \cap \mathcal{S}_2 = \{0\}$ and $x^{\top} y
  = 0$ for all $x \in \mathcal{S}_1, y \in \mathcal{S}_2$. 
  This then generalizes to larger $K$ such that the rows of $X$ are now
  vectors that lie in the union of $K$ orthogonal subspaces
  $\mathcal{S}_1 \cup \mathcal{S}_2 \cup \dots \mathcal{S}_K$ where each
  $\mathcal{S}_k$ is a $K$-dimensional subspace of $\mathbb{R}^{K^2}$,
  and $\Pi$ is a permutation matrix with 
  $K$ fixed points and $K (K - 1)$ cycles of order 2. The permutation
  matrix $\Pi$ then has $K(K-1)/2$ eigenvalues equal to $-1$ (due
  to the $K(K-1)$ cycles of order $2$) and $K(K+1)/2$ eigenvalues equal
  to $1$ (due to the $K(K-1)$ cycles of order $2$ together with the $K$
  fixed points). We can thus write $\Pi = U I_{K(K+1)/2,K(K-1)/2}
  U^{\top}$. We have revised the paper to more clearly present the above
  observations. See also Example~1 on page 9 of the revised manuscript
  and the response to comment $10$ below.

@@

  p8, line 14-21. I don't see why the permutation given by \(\Pi\) has
  \(K\) fixed points.\\
  \textcolor{blue}{
  Example 1 illustrates these fixed points explicitly for $K = 3$. 
  More specifically, comparing the matrices $X$ and $Y$, we see that
  they have three columns in common: the first, fifth, and ninth. 
  The other columns are permuted. The proof of Theorem 1 then
  generalizes this to arbitrary $K$. In particular by the construction
  of $X$, each column of $X$ has exactly one non-zero block
  corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
  points of $\Pi$ are then the columns with indices $r(K+1) +1 $ for $0 \leq r
  \leq K-1$; these columns contain the vectors $\lambda_{kk}$ for
  $1 \leq k \leq K$. The cycles of order $2$ swap the columns of $X$
  containing the vectors $\lambda_{k \ell}$ and $\lambda_{\ell k}$ for
  the $\tbinom{K}{2}$ pairs $(k,\ell)$ with $k \not = \ell$. 

@@

  estimating \(P\) are recalled. A sentence in the beginning of this
  section to expose what is presented could help the reader.\\
  \textcolor{blue}{
  We thank the referee for pointing out this issue. We had revised the
  manuscript to include the line "In our methods, the data that are
  observed is only the adjacency matrix $A \sim
  \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$ along with an assumed
  number of communities, $K$. To motivate our methods, we first
  consider community detection and parameter estimation in the case
  where we know the edge probability matrix, $P$, beforehand, noting
  that community memberships and popularity parameters are not
  immediately discernible from $P$ itself. After establishing methods
  for community detection and parameter estimation from $P$, we use the
  consistency property of ASE to demonstrate that the same methods work
  for $A$ almost surely as $n \to \infty$." Indeed, the assumption of
  known $P$ helps motivate Algorithm $1$ through Algorithm~3 and the
  theoretical results in Theorem~2. Theorem 3 and Theorem 4 then follows
  by leveraging the consistency properties for adjacency spectral embedding
  to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
  community recover and paraemters estimation using the true but unknown $P$. 

@@

\item
  p10. I didn't understand the exposition of the SSC algorithm.\\
  \textcolor{blue}{
  We have added more details to our description of SSC to
  provide further intuition/motivation behind the
  algorithm. The description of SSC now reads:

@@

  construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
  exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
  consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
  The intuition here is that vectors that lie on the same subspace can
  be described as linear combinations of each other, assuming the number
  of vectors in the subspace is greater than the dimensionality of the
  subspace. Thus, for each $c_i$, $c_i^{(j)}$ is zero whenever $x_i$ and $x_j$ belong to different subspaces and may be nonzero otherwise. If \(X\) instead represents points near \(K\) subspaces with some noise, 
  then this property will only hold approximately and 

@@

  p11, line 31. ``\ldots{} SSC on the spectral embedding of \(A\)''.
  Could you support this assertion by references?\\
  \textcolor{blue}{
  There are no references for this, as it is, to the best of our knowledge, a novel idea. 
  We emphasize that while SSC has been proposed for community detection
  for PABM (see e.g., \citep{noroozi2019estimation}), these work apply SSC
  directly on the rows of $A$ (which are binary vectors in
  $\mathbb{R}^{n}$) rather than on the ASE of $A$ (which are real-valued
  vectors in $\mathbb{R}^{d}$ where $d \ll n$). The motivation behind
  our use of SSC comes  from Theorem 1 wherein we show that the latent
  vectors for PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
  by the consistency properties of ASE, the rows of the estimated
  eigenvectors $\hat{V}$ also lies close to the union of orthogonal
  subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
  union of $K$ orthogonal subspaces in $\mathbb{R}^{n}$, the rows of $A$
  (which are now noisy binary vectors) could be quite far from the
  corresponding rows of $P$ and hence need not lie close to the union
  of any $K$ orthogonal subspaces in $\mathbb{R}^{n}$ 

@@

  p14. Do you think that the OSC algorithm can retrieve other kinds of
  structure than assortative ones?\\
  \textcolor{blue}{
  The OSC algorithm works for any PABM parameters setting and is thus
  agnostic to the particular structure (such as assortativity vs
  disassortativity). We had revised the manuscript to include an
  additional simulation demonstrating this claim (see Section 4.3). In
  particular Section~4.3 considers a PABM setting wherein
  $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
  are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$ is a
  vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
  random variables. This result in an average within-community edge
  probability of $1/3$ and a between-community edge
  probability of $2/3$. The results for this simultation setting are
  very similar to those presented in the original manuscript and is also
  consistent with the theoretical results in the paper.

@@

  }\\
  \textcolor{blue}{
  On a related note, it is our view that the distinction between
  assortative and disassortative is not always meaningful for the
  PABM. Indeed, unlike the SBM, in the PABM each vertex is free to have a higher affinity to its
  own community or to other communities. For example, suppose vertex $i$
  belongs to community $1$. Then the average probability of an edge
  between vertex $i$ and another vertex in community $1$ is $n^{-1}
  \rho_n 
  \lambda_{i1} \sum_{j} \lambda_{j1}$ while the average probability of an edge between vertex $i$ and another vertex in community $k \not =
  1$ is $n^{-1} \rho_n \lambda_{ik} \sum_{j} \lambda_{j1}$ and thus, on
  average, vertex $i$ have more affinity with vertices in community $k \not =
  1$ whenever $\lambda_{i1} < \lambda_{ik}$

@@

  misclassification rate to assess the recovery of the nodes
  clustering?\\
  \textcolor{blue}{
  The first two datasets that we analyzed in section 5 had also been analyzed
  in previous papers on the PABM. In particular the
  Leeds Butterfly dataset was analyzed in \citep{noroozi2019estimation}
  and while \citep{noroozi2019estimation} provides enough details
  for us to preprocess the data to match their analysis, they do not
  provide code to run their clustering algorithm (which is based on SSC
  applied to the rows of $A$). Therefore we used their reported
  benchmark values, which are given in terms of ARI. 
  In order to make an apples-to-apples comparison, we also report the ARI for OSC and SSC-ASE here as well. 
  For all other analyses (simulation and real data), we use the
  misclassification error rate/count. 

@@

\item
  Do you consider sparsity in the simulated adjacency matrices?\\
  \textcolor{blue}{
  We have added another simulation in the supplemental materials. 
  In this simulation, we fix $n = 2048$ and $K = 3$ and vary $\rho \in (0, 1)$. 

@@

  extent does the value of \(K\) impact the computational burden for the
  algorithms?\\
  \textcolor{blue}{
  As we discussed in the response to comment
  $14$, we have added a third simulation to address the issue of disassortativity (see Section
  4.3 of the revised manuscript). The performance of all the algorithm
  for the setting in Section~4.3 is qualitatively similar to those
  presented earlier in Section~4.1 and Section~4.2 and is consistent with the theory. 
  Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$. Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 

@@

  }\\
  \textcolor{blue}{
  As for the computational complexity of the algorithms, the main bottleneck for OSC and
  SSC-ASE is the construction of the affinity matrix $B$. 
  For OSC, this involves a (truncated) spectral decomposition to extract the
  $K^2$ largest eigenvalues in modulus (and their corresponding
  eigenvectors) followed by a matrix multiplication of $n \times K^2$
  and $K^2 \times n$ matrices, so the time complexity is of the order $O(n^2 K^2)$. 
  For SSC-ASE (applying SSC on the rows of the adjacency spectral
  embedding $\hat{V}$) we again start with spectral decomposition, followed by
  $n$ LASSO problems with a design matrix of size 
  $K^2 \times (n - 1)$, so the complexity is $O(n^2 K^2 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}. 
  SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
  When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^2 K^2)$. 
  In practice the choice of parameter $K$ does not affect
  runtimes too much. For the simulation studies we only considered $K
  \leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
  comparable estimation accuracy as $K \leq 4$. 

@@

  Again, in your application on the Leeds butterfly dataset, why did you
  limit your analysis to \(K = 4\)?\\
  \textcolor{blue}{
  For the Leeds butterfly dataset, we wanted to compare our results
  using OSC and SSC-ASE to an earlier analysis using SSC on the rows of
  $A$ (see \citep{noroozi2019estimation}). In their analysis they
  describe how they removed some of the nodes (corresponding to the
  other butterfly species) so that the resulting dataset only include the
  $K = 4$ butterfly species as considered in our paper.}

@@

  }\\
  \textcolor{blue}{
  Therefore, the dimensionality of the latent configuration is not in question, only what kind of effect multiplication by unidentifiable $Q$ has on the configuration. 
  The OSC algorithm circumvents the nonidentifiability of the latent
  configuration by only considering the entries of $\hat{B} = n \hat{V}
  \hat{V}^{\top}$ as $n\hat{V}\hat{V}^{\top}$ is, by the asymptotic properties of
  adjacency spectral embedding, close to $n VV^{\top}$ and $n V
  V^{\top}$ is always unique (the orthogonal projection onto a column
  space is always unique). Similarly, SSC-ASE also circumvents this non-identifiability
  by only using the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
  $\hat{v}_j$ of $\hat{V}$ belonging to different subspaces, i.e.,
  different communities, will also
  have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 

@@

  }\\
  \textcolor{blue}{
  That said, looking back at the latent configuration in Theorem 1, we
  can see that a $K$-dimensional (not $K^2$-dimensional) embedding is
  sufficient for clustering, i.e., the embedding corresponding to the
  diagonal blocks $\lambda_{11}, \lambda_{22}, \dots, \lambda_{KK}$. 
  If we are able to find this embedding then this would greatly reduce
  the number of dimensions required for the community
  recovery/clustering. In practice, this isn't always possible due to
  the nonidentifiability issue; in particular, determining which of the
  $K$ out of $K^2$ eigenvectors should be used is the main difficulty.
  Nevertheless we have found via numerical experiments that community
  detection is often possible using only the eigenvectors corresponding
  to the $2^\mathrm{nd}$ to the $(K+1)^\mathrm{st}$ largest eigenvalues
  of $A$. 

@@

  understand why this has an impact on the performance for \(K = 2\),
  but not for \(K \geq 3\).\\
  \textcolor{blue}{
  We are not entirely clear on why this happens for $K = 2$, but when
  comparing the normalized Laplacian eigenmaps of the affinity matrix
  $B$ for various $K$, we see that when $K > 2$, the points appear much
  closer to Gaussian mixtures, whereas with $K = 2$, while there is some
  separation by community, they do not appear as Gaussian mixtures. 
  One possible reason for this may be because we failed to find the
  right value for the hyperparameter $\vartheta$ for SSC-ASE. 

@@

  }\\
  \textcolor{blue}{
  More specifically, under ideal circumstances, we would be able to
  circumvent the issue of selecting the hyperparameter $\vartheta$ by
  using the theoretical value that ensures, with high probability, the rows of $\hat{V}$ satisfy the subspace
  detection property (SDP). However, as other researchers had pointed out (see
  e.g., \citep{sdp_sufficiency,liu_ssc}, the fact that SDP is satisfied
  does not guarantee that the affinity matrix $B$ represents a graph
  with $K$ connected components but rather a graph with {\em at least}
  $K$ connected component. Indeed, we found in our simulations that
  setting $\vartheta$ to the theoretical values that guarantees 
  SDP often results in $B$ having more than $K$ connected components.
  This is why we choose to do spectral clustering on $B$, i.e., we first
  compute the normalized Laplacian embeddings using $B$ before
  clustering the rows of this embedding using Gaussian mixtures
  modeling. 

@@

  }\\
  \textcolor{blue}{
  Finally we note that in our simulations the proportion of mislabeled
  vertices does decreases as $n$ increases. 
  For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
  vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
  $n = 4096$; this correspond to an error rate of around $16\%$ and
  $2\%$, respectively.

@@

  reason for the different behavior?\\
  \textcolor{blue}{
  In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 
  While all three algorithms are specific to the PABM, they might behave
  differently to the different ways in which the model is
  misspecified. Finally, we note that Section~5 analyzed three real
  datasets. For the first dataset (Leeds butterfly) we didn't run the MM
  algorithm as the original implementation of MM used in 
  \cite{307cbeb9b1be48299388437423d94bf1} is extremely slow when $K = 4$. For
  the second dataset (British MP's), the accuracy of MM is slightly
  better than that of OSC. Finally, for the last dataset (Karantaka
  villages), the accuracy of MM is slightly worse than that of OSC. }

### 2022-02-11 summary_minh_edited.tex
minor update referencing example 1 in proof of theorem 1
Premerge
minor typos

For the general case, we can extend $\tilde{P} = X \Pi X^\top$ to larger $K$. 
We once again consider $\tilde{P}$ as defined in
Remark~\ref{rem:pabm_view2}.  
We first define the following matrices
\begin{gather}
\label{eq:xy}
\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k1)} \mid \cdots \mid \lambda^{(kK)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}, \quad
X = \mathrm{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
\mathbb{R}^{n \times K^2}, \\
L^{(k)} = \mathrm{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
\mathbb{R}^{n \times K}, \quad
Y = \begin{bmatrix} L^{(1)} \mid \cdots \mid L^{(K)} \end{bmatrix} \in
\mathbb{R}^{n \times K^2}.
\end{gather}
It is then straightforward to verify that
\begin{gather*}
  XY^{\top} = \mathrm{blockdiag}(\Lambda^{(1)}, \dots,
\Lambda^{(K)}) \begin{bmatrix} L_1^{\top} \\ \vdots \\
  L_{K}^{\top} \end{bmatrix} = \begin{bmatrix} \Lambda^{(1)}
  (L^{(1)})^{\top} \\ \vdots \\
  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix}, \\
\Lambda^{(k)} (L^{(k)})^{\top} = \begin{bmatrix} \lambda^{(k1)}
  (\lambda^{(1k)})^{\top} \mid \dots \mid \lambda^{(kK)}
  (\lambda^{(Kk)})^{\top} \end{bmatrix} = \begin{bmatrix} P^{(k1)}
  \mid P^{(k2)} \mid \dots \mid P^{(kK)} \end{bmatrix}.
\end{gather*}
We therefore have $\tilde{P} = X Y^\top$. 
Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
$\Pi$ and hence $\tilde{P} = X \Pi X^\top$.
The permutation described by $\Pi$ now has $K$ fixed points, which correspond to
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where
$k = r (K + 1) + 1$ for $0 \leq r \leq K - 1$. It also has
$\tbinom{K}{2}$ cycles of order $2$. Each cycle corresponds to
a pair of eigenvalues $\{-1,+1\}$ and a pair of eigenvectors
$\{(e_s + e_t)/\sqrt{2},(e_s - e_t)/ \sqrt{2}\}$.

@@

In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself. 
After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.

@@

In practice, due to presence of noise, SSC is often done by solving the LASSO problems

\begin{equation} \label{eq:ssc}
c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i}^\top c\|^2_2 + \vartheta \|c\|_1
\end{equation}

for some sparsity parameter \(\vartheta > 0\). 
The \(c_i\) vectors are then collected into \(C\) and \(B\) as before.

@@

We previously stated in Theorem~\ref{theorem2} one possible set of latent positions that result in
the edge probability matrix of a PABM, namely
\(P = \tilde{\Pi} (XU) I_{p, q}
(XU)^\top \tilde{\Pi}^{\top}\) where $X$ is block diagonal and
$\tilde{\Pi}$ is a permutation matrix.  
Furthermore, the explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
such that points within each community lie on \(K\)-dimensional
orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $i$ and $j$ are in different communities. 
Thus if we have (or can estimate) \(XU\) directly, then both the community
detection and parameter identification problem are trivial because \(U\)
is orthonormal and fixed for each value of \(K\). 
However, direct
identification or estimation of \(XU\) is possibly difficult 
due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable})
when we are given only $P$. 
More specifically, suppose we find a matrix $Y \in
\mathbb{R}^{n \times K^2}$
such that \(P = Y I_{p, q} Y^\top\). Then it is generally the case that
we  \( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
$Q \in \mathbb{O}(p,q)$. 
However since \(Q\) is not necessarily an
orthogonal matrix and hence, if $y_i$ denote the $i^{th}$ row of $Y$, 
then $\langle U^{\top} x_i, U^{\top} x_j \rangle
\neq \langle y_i, y_j \rangle$.
This prevents us from transferring the orthogonality property of
\(XU\) directly to $Y$.

>>>>

For the general case, we can extend $\tilde{P} = X \Pi X^\top$ to larger $K$. 
For a more concrete example of this, refer to Example~1. 
We once again consider $\tilde{P}$ as defined in
Remark~\ref{rem:pabm_view2}.  
We first define the following matrices

\begin{gather}
\label{eq:xy}
\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k1)} \mid \cdots \mid \lambda^{(kK)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}, \quad
X = \mathrm{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
\mathbb{R}^{n \times K^2}, \\
L^{(k)} = \mathrm{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
\mathbb{R}^{n \times K}, \quad
Y = \begin{bmatrix} L^{(1)} \mid \cdots \mid L^{(K)} \end{bmatrix} \in
\mathbb{R}^{n \times K^2}.
\end{gather}

It is then straightforward to verify that

\begin{gather*}
  XY^{\top} = \mathrm{blockdiag}(\Lambda^{(1)}, \dots,
\Lambda^{(K)}) \begin{bmatrix} L_1^{\top} \\ \vdots \\
  L_{K}^{\top} \end{bmatrix} = \begin{bmatrix} \Lambda^{(1)}
  (L^{(1)})^{\top} \\ \vdots \\
  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix}, \\
\Lambda^{(k)} (L^{(k)})^{\top} = \begin{bmatrix} \lambda^{(k1)}
  (\lambda^{(1k)})^{\top} \mid \dots \mid \lambda^{(kK)}
  (\lambda^{(Kk)})^{\top} \end{bmatrix} = \begin{bmatrix} P^{(k1)}
  \mid P^{(k2)} \mid \dots \mid P^{(kK)} \end{bmatrix}.
\end{gather*}
We therefore have $\tilde{P} = X Y^\top$. 
Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
$\Pi$ and hence $\tilde{P} = X \Pi X^\top$.
The permutation described by $\Pi$ now has $K$ fixed points, which correspond to
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where
$k = r (K + 1) + 1$ for $0 \leq r \leq K - 1$. It also has
$\tbinom{K}{2}$ cycles of order $2$. Each cycle corresponds to
a pair of eigenvalues $\{-1,+1\}$ and a pair of eigenvectors
$\{(e_s + e_t)/\sqrt{2},(e_s - e_t)/ \sqrt{2}\}$.

@@

In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
To motivate our methods, we first consider community detection and parameter estimation in the case where we know the edge probability matrix $P$ beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself. 
After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.

@@

In practice, due to presence of noise, SSC is often done by solving the LASSO problems
\begin{equation} \label{eq:ssc}
c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i}^\top c\|^2_2 + \vartheta \|c\|_1
\end{equation}
for some sparsity parameter \(\vartheta > 0\). 
The \(c_i\) vectors are then collected into \(C\) and \(B\) as before.

@@

We previously stated in Theorem~\ref{theorem2} one possible set of latent positions that result in
the edge probability matrix of a PABM, namely
\(P = \tilde{\Pi} (XU) I_{p, q}
(XU)^\top \tilde{\Pi}^{\top}\) where $X$ is block diagonal and
$\tilde{\Pi}$ is a permutation matrix.  
Furthermore, the explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
such that points within each community lie on \(K\)-dimensional
orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $i$ and $j$ are in different communities. 
Thus if we have (or can estimate) \(XU\) directly, then both the community
detection and parameter identification problem are trivial because \(U\)
is orthonormal and fixed for each value of \(K\). 
However, direct
identification or estimation of \(XU\) is possibly difficult 
due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable})
when we are given only $P$. 
More specifically, suppose we find a matrix $Y \in
\mathbb{R}^{n \times K^2}$
such that \(P = Y I_{p, q} Y^\top\). Then it is generally the case that
\( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
$Q \in \mathbb{O}(p,q)$. 
However since \(Q\) is not necessarily an
orthogonal matrix and hence, if $y_i$ denote the $i^{th}$ row of $Y$, 
then $\langle U^{\top} x_i, U^{\top} x_j \rangle
\neq \langle y_i, y_j \rangle$.
This prevents us from transferring the orthogonality property of
\(XU\) directly to $Y$.
